---
title: "my first knitr"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library("data.table")
library(ggplot2)
```

## Introduction
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, orJawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals throughout the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day

## Loading and preprocessing the data

```{r steps}
## Read data file and understand the summary of the data set

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = paste0(getwd(), '/repdata%2Fdata%2Factivity.zip'), method = "curl")
unzip("repdata%2Fdata%2Factivity.zip",exdir = "data")
activityDT <- data.table::fread(input = "data/activity.csv")

summary(activityDT)


str(activityDT)

## Calculate the total number of steps taken per day

Total_Steps <- activityDT[, c(lapply(.SD, sum, na.rm = FALSE)), .SDcols =c("steps"), by = .(date)] 

head(Total_Steps, 10)

```

## Including Plots

```{r plots, echo=FALSE}
## Make a histogram of the total number of steps taken each day

ggplot(Total_Steps, aes(x = steps)) +
  geom_histogram(fill = "blue", binwidth = 1000) +
  labs(title = "Daily Steps", x = "Steps", y = "Frequency")

## calculate mean and media of steps taken each day

Total_Steps[, .(Mean_Steps = mean(steps, na.rm = TRUE), Median_Steps = median(steps, na.rm = TRUE))]

## time series plot showing averge steps taken for 5-minute interval

IntervalDT <- activityDT[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval)] 

ggplot(IntervalDT, aes(x = interval , y = steps)) + geom_line(color="blue", size=1) + labs(title = "Avg. Daily Steps", x = "Interval", y = "Avg. Steps per day")

## max step for the 5-minute interval

IntervalDT[steps == max(steps), .(max_interval = interval)]

```

## Including Plots - Imputing missing values

```{r plots2, echo=FALSE}
## number of missing values in the data set

activityDT[is.na(steps), .N ]

## filling missing values with median

activityDT[is.na(steps), "steps"] <- activityDT[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]

## create new data set without missing values

data.table::fwrite(x = activityDT, file = "data/tidyData.csv", quote = FALSE)

Total_Steps <- activityDT[, c(lapply(.SD, sum)), .SDcols = c("steps"), by = .(date)] 

## mean and median number of steps taken per day (no NAs)
Total_Steps[, .(Mean_Steps = mean(steps), Median_Steps = median(steps))]

## steps taken a day (no NAs)

ggplot(Total_Steps, aes(x = steps)) + geom_histogram(fill = "blue", binwidth = 1000) + labs(title = "Daily Steps", x = "Steps", y = "Frequency")

## old and new mean and median

mean <- c(10766,9354)
median <- c(10765,10395)
result <- data.frame(mean,median)
rownames(result) <- c("oridata","imputeddata")
result
